### 2024-04-22 paper/paper.qmd
paper and parquets

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Jessica Im
thanks: "Code and data are available at: https://github.com/jimessica/boycott"
date: today
date-format: long
abstract: "Introduction to the topic and its importance. What I did in my research and analysis. What was found from my research and analysis. Significance of my findings."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tinytex)
library(kableExtra)
library(knitr)
library(janitor)
library(kableExtra)
library(ggplot2)
library(xts)
library(tidyquant)
library(arrow)

cleaned_data_NKE <- read.csv("../data/analysis_data/cleaned_data_NKE.csv")
cleaned_data_NKE <- 
  cleaned_data_NKE |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_NKE$Date <- 
  as.Date(cleaned_data_NKE$Date, format='%Y-%m-%d')

cleaned_data_PRKS <- read.csv("../data/analysis_data/cleaned_data_PRKS.csv")
cleaned_data_PRKS <- 
  cleaned_data_PRKS |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_PRKS$Date <- 
  as.Date(cleaned_data_PRKS$Date, format='%Y-%m-%d')

cleaned_data_SBUX <- read.csv("../data/analysis_data/cleaned_data_SBUX.csv")
cleaned_data_SBUX <- 
  cleaned_data_SBUX |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_SBUX$Date <- 
  as.Date(cleaned_data_SBUX$Date, format='%Y-%m-%d')

raw_data_SPX <- tq_get('SPX',
               from = "1991-07-12",
               to = "2024-03-29",
               get = "stock.prices")

cleaned_data_SPX <-
  raw_data_SPX |>
  select(symbol, date, close)

cleaned_data_DJI <-read_parquet("../data/analysis_data/cleaned_data_DJI.parquet")

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

In [Background], the history of why each company was or is being boycotted is explained. In the following section, [Data],

# Background

## Nike (NKE)

In August of 1992, Jeffrey Ballinger published “The new free-trade heel: Nike's profits jump on the backs of Asian workers” in a Harper’s Bazaar issue [@harpers]. This publication revealed the low pay and poor working conditions in Nike’s Southeast Asian and Middle Eastern sweatshops [@ethically]. Later that year, Nike released the “Code of Conduct and Memorandum of Understanding” for contract workers, but the labour controversy was well underway. In the following years, protests in Southeast Asia and the United States raged on. Protests at the Barcelona Olympics in 1992 and the founding of Ballinger’s labour group Press for Change in 1993 brought new waves of media attention to the cause. After years of Nike promotions fueling more public outrage, in 1998, Nike’s then-CEO Phil Knight announced that Nike will raise minimum wage, increase monitoring, and introduce U.S. clean air standards at all factories [@ethically]. Knight acknowledged the brand’s association with workplace abuse and stated that "the American consumer doesn’t want to buy products made under abusive conditions”. This speech marked the shift away from boycotting and for the purposes of this paper will represent the end of Nike boycotting.

## United Parks and Recreation
Blackfish is a film focused on the cruel captivity of orcas by focusing on the infamous killer whale, Tilikum, who was kept in captivity at SeaWorld and was involved in the death of 3 people [@WAP]. The film was released in 2013 and shown on major channels like CNN, Netflix, and Universal Studios around the world and over its 14-week box office period in North America, earned $2,073,582 [@wikiBlackfish]. In March of 2014, California Assemblyman Richard Bloom proposed that orcas are banned from being held captive for entertainment [@CNNorcaban]. Later that year, U.S. Congressmen Adam Schiff and Jared Huffman amended the Agriculture Appropriations Act to require the USDA’s Animal & Plant Health Inspection Service to update and finalize regulations on marine mammal captivity [@SchiffandHuffman]. On behalf of SeaWorld investors, the Rosen Law Firm filed a class action lawsuit against SeaWorld in 2014 for

## Starbucks
On October 9th the Starbucks Workers United posted to their Twitter “Solidarity with Palestine” [@apnews]. This tweet was up for 40 minutes before it was taken down and was originally posted without the approval of union leaders according to Workers United. Even after the post was taken down, posts and retweets from local Starbucks United twitter accounts in support of Palestine and condemning Israel were circulating. On October 13th, Starbucks sent a letter to Workers United, demanding the union to stop using the Starbucks name and similar logo. Workers United retaliated saying that Starbucks was using the tragedy in the Middle East to support their anti-union campaign.

# Data {#sec-data}
A sense of the dataset should be communicated to the reader. The broader context of the dataset should be discussed. All variables should be thoroughly examined and explained. Explain if there were similar datasets that could have been used and why they were not. If variables were constructed then this should be mentioned, and high-level cleaning aspects of note should be mentioned, but this section should focus on the destination, not the journey. It is important to understand what the variables look like by including graphs, and possibly tables, of all observations, along with discussion of those graphs and the other features of these data. Summary statistics should also be included, and well as any relationships between the variables. If this becomes too detailed, then appendices could be used. Basically, for every variable in your dataset that is of interest to your paper there needs to be graphs and explanation and maybe tables.

## Source Data and Methodology
The data used in this paper is from 
We want the reader to understand what the data that underpin the results look like.

Stock prices go up when people think it's worth more than what it is trading at -> supply and demand -> as more people want it the prices goes up.

## Limitations

## Data Cleaning
In @tbl-rawdataNKE, the data's date and stock price are shown.

```{r}
#| label: tbl-rawdataNKE
#| tbl-cap: "Description of Raw Nike Stock Variables"
#| echo: false

rawdataNKE <- data.frame( 
  "Variables" = c( "Date", "Close"),
  "New Name" = c( "Date", "Stock Price"), 
  "Description of Variables" = c("The day of the stock price", "The value of the stock at market close"), 
  "Example Value" = c("1991-07-12", "1.355469")
)

knitr::kable(rawdataNKE, 
      col.names = c("Variables", 
                    "New Name",
                    "Description of Variables" ,
                    "Example Value"),
      booktabs = TRUE,
      linesep = " "
      )
```

## Stock Prices

@fig-linegraphNKE shows the change in stock price during boycotting of Nike.
```{r}
#| label: fig-linegraphNKE
#| fig-cap: Changes in Nike Stock Price from One Week Before the Commencement of Labour Boycotts to One Week After Their Pledge to End Poor Working Conditions and Child Labour
#| echo: false

cleaned_data_NKE |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "July 12th, 1991 to May 19th, 1998", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "12 months",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(0,10) + 
  theme_minimal()

```

@fig-linegraphPRKS shows the change in stock price during boycotting of United Parks and Recreation.
```{r}
#| label: fig-linegraphPRKS
#| fig-cap: Changes in United Parks and Recreation Stock Price from One Week Before Blackfish was Released to One Week After Their Pledge to End Orca Breeding
#| echo: false

cleaned_data_PRKS |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "July 12th, 2013 to May 17th, 2016", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "12 months",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(15,40) + 
  theme_minimal()

```

@fig-linegraphsSBUX shows 
```{r}
#| label: fig-linegraphSBUX
#| fig-cap: Changes in Starbucks Stock Price from One Week Before Legal Action Against Workers United to March 28th
#| echo: false

cleaned_data_SBUX |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "July 12th, 2013 to May 17th, 2016", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "1 month",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(80,110) + 
  theme_minimal()
```

\newpage

# Results

## Overall Change in Stock Price at Commencement and End of Boycott
@tbl-starttoendNKE, @tbl-starttoendPRKS, and @tbl-starttoendSBUX summarizes the change in stock price between a week before the boycott commenced and its commencement date as well as the difference between the start and end dates of the boycott. The first date is the stock price a week before the stock price and the second date is the day the boycott began.

@tbl-starttoendNKE shows a small decrease in stock price from a week before the beginning of the boycott. There is also an increase in stock price from the day the boycott started to the day it ended. Relative to the stock price, the change is very large as it multiplied by 3.34 times. 
```{r}
#| label: tbl-starttoendNKE
#| tbl-cap: Changes in Starbucks Stock Price from One Week Before Legal Action Against Workers United, the the Day it was Announced, to March 28th
#| echo: false

starttoend_NKE <- cleaned_data_NKE |>
  filter(Date == '1991-07-12' | Date == '1991-07-18' | Date == '1998-05-19')

starttoend_NKE <-
  starttoend_NKE |> 
  mutate(Difference = Close - lag(Close))

kable(starttoend_NKE, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference Between Stock Prices"),
      booktabs = TRUE,
      linesep = ""
)
```

@tbl-starttoendPRKS shows that there was a significant decrease in stock price from the beginning to end of the boycott. 
```{r}
#| label: tbl-starttoendPRKS
#| tbl-cap: Changes in United Parks & Recreation Stock Price from One Week Before Blackfish was Released, to the Date it was Released, to their Announcement that they will end their Orca Breeding Program
#| echo: false

starttoendPRKS <- cleaned_data_PRKS |>
  filter(Date == '2013-07-12' | Date == '2013-07-19' | Date == '2016-03-16')

starttoendPRKS <-
  starttoendPRKS |> 
  mutate(Difference = Close - lag(Close))

kable(starttoendPRKS, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference Between Stock Prices"),
      booktabs = TRUE,
      linesep = ""
)
```

@tbl-starttoendSBUX shows minimal change in stock price throughout the boycott. 
```{r}
#| label: tbl-starttoendSBUX
#| tbl-cap: Changes in Starbucks Stock Price from One Week Before Legal Action Against Workers United, the the Day it was Announced, to March 28th
#| echo: false

starttoendSBUX <- cleaned_data_SBUX |>
  filter(Date == '2023-10-11' | Date == '2023-10-18' | Date == '2024-03-28')

starttoendSBUX <-
  starttoendSBUX |> 
  mutate(Difference = Close - lag(Close))

kable(starttoendSBUX, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference Between Stock Prices"),
      booktabs = TRUE,
      linesep = ""
)
```

@tbl-twoweeksSBUX

```{r}
#| label: tbl-twoweeksSBUX
#| tbl-cap: Changes in Starbucks Stock Price Two Weeks From the Lawsuit
#| echo: false

twoweeksSBUX <- 
  cleaned_data_SBUX |>
  filter(between(Date, as.Date('2023-10-18'), as.Date('2023-11-06')))

twoweeksSBUX <-
  twoweeksSBUX |> 
  mutate(Difference = Close - lag(Close))

kable(twoweeksSBUX, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference"),
      booktabs = TRUE,
      linesep = ""
)
```

@tbl-SPXandPRKS
```{r}
#| label: tbl-SPXandPRKS
#| tbl-cap: Changes in Starbucks Stock Price from One Week Before Legal Action Against Workers United, the the Day it was Announced, to March 28th
#| echo: false

PRKS_SPX <-
  cleaned_data_SPX |>
  filter(between(date, as.Date('2013-07-19'), as.Date('2016-03-16')))

PRKS_Focus <-
  cleaned_data_PRKS |>
  filter(between(Date, as.Date('2013-07-19'), as.Date('2016-03-16')))

SPX_Stocks <- PRKS_SPX$close
PRKS_Stocks <- PRKS_Focus$Close

PRKS_Date <- PRKS_Focus$Date

par(mfrow=c(2,1))
par(mar = c(0, 0, 0, 0))
plot(PRKS_Date,PRKS_Stocks, type="l", col="#0000c2" )
plot(PRKS_Date,SPX_Stocks, type="l", col="#FF5733" )

```

@tbl-DJIandNIKE
```{r}
#| label: tbl-DJIandNIKE
#| tbl-cap: Changes in Starbucks Stock Price from One Week Before Legal Action Against Workers United, the the Day it was Announced, to March 28th
#| echo: false

NIKE_DJI <-
  cleaned_data_DJI |>
  filter(between(date, as.Date('1991-07-18'), as.Date('1998-05-19')))

NIKE_Focus <-
  cleaned_data_NKE |>
  filter(between(Date, as.Date('1991-07-18'), as.Date('1998-05-19')))

DJI_Stocks <- NIKE_DJI$close
NIKE_Stocks <- NIKE_Focus$Close

NIKE_Date <- NIKE_Focus$Date

par(mfrow=c(2,1))
par(mar = c(0, 0, 0, 0))
plot(NIKE_Date,NIKE_Stocks, type="l", col="#0000c2" )
plot(NIKE_Date,DJI_Stocks, type="l", col="#FF5733" )

```

# Discussion

## Out of Sight, Out of Mind: Nike isn't the Only One

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## SeaWorld: The Tainted Consumer Experience

## Starbucks: Impact of Unionization on Stockprice
Investors care about the company's cash flows/earnings margin, when unions are created, the employees get better treatment the earnings margin is affects and is lowered, which investors do not like. Thus, it is likely that stock price went up when Starbucks and Workers United decided to sue each other. They think that Starbucks is gonna win the lawsuit,

On November 2nd, Starbucks released their Q4 and Full Year Fiscal Results. Some highlights were record net revenues, store sales up by 8% globally, and record number of stores [@investstarbucks]

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References