### 2024-04-26 paper/paper.qmd
updating to add NDX and add paper writing

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Jessica Im
thanks: "Code and data are available at: [https://github.com/jimessica/boycott)](https://github.com/jimessica/boycott)"
date: today
date-format: long
abstract: "Introduction to the topic and its importance. What I did in my research and analysis. What was found from my research and analysis. Significance of my findings."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tinytex)
library(kableExtra)
library(knitr)
library(janitor)
library(kableExtra)
library(ggplot2)
library(xts)
library(tidyquant)
library(arrow)
library(readr)
library(gridExtra)

## Import cleaned data and turn "Date" into class Date
# Nike
cleaned_data_NKE <- read.csv("../data/analysis_data/cleaned_data_NKE.csv")
cleaned_data_NKE <- 
  cleaned_data_NKE |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_NKE$Date <- 
  as.Date(cleaned_data_NKE$Date, format='%Y-%m-%d')

# United Parks and Recreation
cleaned_data_PRKS <- read.csv("../data/analysis_data/cleaned_data_PRKS.csv")
cleaned_data_PRKS <- 
  cleaned_data_PRKS |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_PRKS$Date <- 
  as.Date(cleaned_data_PRKS$Date, format='%Y-%m-%d')

#Starbucks
cleaned_data_SBUX <- read.csv("../data/analysis_data/cleaned_data_SBUX.csv")
cleaned_data_SBUX <- 
  cleaned_data_SBUX |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_SBUX$Date <- 
  as.Date(cleaned_data_SBUX$Date, format='%Y-%m-%d')

## Import S&P 500 Index and Dow Jones Index Data
# S&P 500
cleaned_data_SPX <-read_parquet("../data/analysis_data/cleaned_data_SPX.parquet")

# Dow Jones 
cleaned_data_DJI <-read_parquet("../data/analysis_data/cleaned_data_DJI.parquet")

# Nasdaq 100
cleaned_data_NDX <-read_parquet("../data/analysis_data/cleaned_data_NDX.parquet")

```

# Introduction
Boycotting is a common way for consumers to cast their vote on the amount of power corporations and institutions should have. Boycotting is usually done by avoiding purchasing from a company or withdrawing support for an institution [@ethicalconsumer]. It is oftentimes a response to an ethical violation and has had most success by impacting the public image of the company through generating negative media coverage [@northwestern]. As businesses have grown and markets have been dominated by oligopolies, consumers feel ignored and alienated [@kokuthreats]. As a result, boycotting has increased and between 1984 to 1990 boycotting increased by 769%. Recently, the Boycott, Divestment, Sanctions (BDS) movement in support of Palestine’s freedom and justice has introduced a new wave of companies being boycotted. Many debates surround the effectiveness of boycotting and whether it can inflict substantial financial losses for  targeted organizations.

The objective of this paper is to determine the impact of boycotting on stock prices and the role earnings calls, product or service type, and social media has. To investigate, Nike, United Parks and Recreation, and Starbucks’s stock prices will be analyzed.

In the following section, [Background](@sec-background), the history of why each company was or is being boycotted is explained. In [Data](#sec-data), the dataset, variables used, limitations of the data, and methodology for analysis and cleaning are explained. In [Results](#sec-results), there are detailed explorations of the difference in stock price over the boycott, comparisons with the overall market, and an analysis of important dates during the boycotts. Finally, the [Discussion](#sec-discussion) explores the impact of earnings calls, product or service type, and social media.

# Background {#sec-background}

## Nike (NKE)

In August of 1992, Jeffrey Ballinger published “The new free-trade heel: Nike's profits jump on the backs of Asian workers” in a Harper’s Bazaar issue [@harpers]. This publication revealed the low pay and poor working conditions in Nike’s Southeast Asian and Middle Eastern sweatshops [@ethically]. Later that year, Nike released the “Code of Conduct and Memorandum of Understanding” for contract workers, but the labour controversy was well underway. In the following years, protests in Southeast Asia and the United States raged on. Protests at the Barcelona Olympics in 1992 and the founding of Ballinger’s labour group Press for Change in 1993 brought new waves of media attention to the cause. After years of Nike promotions fueling more public outrage, in 1998, Nike’s then-CEO Phil Knight announced that Nike will raise minimum wage, increase monitoring, and introduce U.S. clean air standards at all factories [@ethically]. Knight acknowledged the brand’s association with workplace abuse and stated that "the American consumer doesn’t want to buy products made under abusive conditions”. This speech marked the shift away from boycotting and for the purposes of this paper will represent the end of Nike boycotting.

## United Parks and Recreation

Blackfish is a film focused on the cruel captivity of orcas by focusing on the infamous killer whale, Tilikum, who was kept in captivity at SeaWorld and was involved in the death of 3 people [@WAP]. The film was released in 2013 and shown on major channels like CNN, Netflix, and Universal Studios around the world and over its 14-week box office period in North America, earned \$2,073,582 [@wikiBlackfish]. In March of 2014, California Assemblyman Richard Bloom proposed that orcas are banned from being held captive for entertainment [@CNNorcaban]. Later that year, U.S. Congressmen Adam Schiff and Jared Huffman amended the Agriculture Appropriations Act to require the USDA’s Animal & Plant Health Inspection Service to update and finalize regulations on marine mammal captivity [@SchiffandHuffman]. On behalf of SeaWorld investors, the Rosen Law Firm filed a class action lawsuit against SeaWorld in 2014 for

## Starbucks

On October 9th the Starbucks Workers United posted to their Twitter “Solidarity with Palestine” [@apnews]. This tweet was up for 40 minutes before it was taken down and was originally posted without the approval of union leaders according to Workers United. Even after the post was taken down, posts and retweets from local Starbucks United twitter accounts in support of Palestine and condemning Israel were circulating. On October 13th, Starbucks sent a letter to Workers United, demanding the union to stop using the Starbucks name and similar logo. Workers United retaliated saying that Starbucks was using the tragedy in the Middle East to support their anti-union campaign.

# Data {#sec-data}

A sense of the dataset should be communicated to the reader. The broader context of the dataset should be discussed. All variables should be thoroughly examined and explained. Explain if there were similar datasets that could have been used and why they were not. If variables were constructed then this should be mentioned, and high-level cleaning aspects of note should be mentioned, but this section should focus on the destination, not the journey. It is important to understand what the variables look like by including graphs, and possibly tables, of all observations, along with discussion of those graphs and the other features of these data. Summary statistics should also be included, and well as any relationships between the variables. If this becomes too detailed, then appendices could be used. Basically, for every variable in your dataset that is of interest to your paper there needs to be graphs and explanation and maybe tables.

## Source Data and Methodology

The source data was obtained from Yahoo Finance Canada [@yahoo] and the R package [@R-tidyquant]. The Yahoo Finance Canada historical data was filtered for the desired date range and daily data before being downloaded. The original data had “Date”, “Open”, “High”, “Low”, “Close”, “Adj Close”, and “Volume”. The only variables of interest were “Date” and “Close” which represented the date of the stock’s price and the price when the market closed respectively.

## Limitations

## Data Cleaning
The data was cleaned using the statistical programming language R [@citeR] as well as library packages `tidyverse` [@R-tidyverse], `tinytex` [@R-tinytex],`ggplot2` [@R-ggplot2], `dplyr` [@R-dplyr], `readr` [@R-readr], `tibble` [@R-tibble], `stringr` [@R-stringr], `kableExtra` [@R-kableExtra], `janitor` [@R-janitor],  `xts` [@R-xts], and `knitr` [@knitr2015], and `arrow` [@R-arrow]. The data was cleaned by selecting the two variables “Date” and “Close”  which are shown in @tbl-rawdataNKE below @tbl-rangeData, shows the number of days observed as part of the boycott in addition to one week before it started. It also indicates the start and end date of each dataset. The largest range was Nike at nearly 7 years compared to the less than 4 months Starbucks has been observed for. Each company also represents a different decade, with Nike in the 1990s, Parks in the 2010s, and most recently Starbucks in the 2020s.

```{r}
#| label: tbl-rawdataNKE
#| tbl-cap: "Description of Raw Nike Stock Variables"
#| echo: false

rawdataNKE <- data.frame( 
  "Variables" = c( "Date", "Close"),
  "New Name" = c( "Date", "Stock Price"), 
  "Description of Variables" = c("The day of the stock price", "The value of the stock at market close"), 
  "Example Value" = c("1991-07-12", "1.355469")
)

knitr::kable(rawdataNKE, 
      col.names = c("Variables", 
                    "New Name",
                    "Description of Variables" ,
                    "Example Value"),
      booktabs = TRUE,
      linesep = " "
      )
```

```{r}
#| label: tbl-rangeData
#| tbl-cap: "The range of time each company's stock was observed for."
#| echo: false

N_NKE = nrow(cleaned_data_NKE)
N_PRKS = nrow(cleaned_data_PRKS)
N_SBUX = nrow(cleaned_data_SBUX)

Day1_NKE = cleaned_data_NKE[1,1]
Day2_NKE = cleaned_data_NKE[N_NKE,1]

Day1_PRKS = cleaned_data_PRKS[1,1]
Day2_PRKS = cleaned_data_PRKS[N_PRKS,1]

Day1_SBUX = cleaned_data_SBUX[1,1]
Day2_SBUX = cleaned_data_SBUX[N_SBUX,1]

dataALL <- data.frame(
  "Companies" = c("Nike (NKE)", "United Parks and Recreation (PRKS)", "Starbucks (SBUX)"),
  "Days Observed" = c(N_NKE, N_PRKS, N_SBUX),
  "Date Start" = c(Day1_NKE, Day1_PRKS, Day1_SBUX),
  "Date End" = c(Day2_NKE, Day2_PRKS, Day2_SBUX)
)

knitr::kable(dataALL, 
      col.names = c("Companies", 
                    "Days Observed",
                    "Date Start" ,
                    "Date End"),
      booktabs = TRUE,
      linesep = " "
      )
```

## Change in Stock Price
The change in stock price over the length of the boycott is observed in @fig-linegraphNKE, @fig-linegraphPRKS, and @fig-linegraphSBUX. These three graphs show a week before the boycott started to the end of the boycott. @fig-linegraphNKE shows how Nike’s stock price was relatively unstable in the first three years of the boycott, with stock prices rising later in 1991 before dropping slightly and then peaking early in 1993 before descending to a low a year later. A progressive cimb from that point to February of 1997 sees a growth of around 750%. @fig-linegraphPRKS shows that United Parks and Recreation had a steep drop in stock price when ‘Blackfish’ released and had a year of poor stability after before dropping drastically. Afterwards the stock price continues to remain unstable.  @fig-linegraphSBUX graphs the stock price of Starbucks and depicts a steep climb in price in November of 2023 with a peak in early December before a steady climb down over the next 5 months.

```{r}
#| label: fig-linegraphNKE
#| fig-cap: Changes in Nike Stock Price from One Week Before the Commencement of Labour Boycotts to One Week After Their Pledge to End Poor Working Conditions and Child Labour
#| echo: false

cleaned_data_NKE |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "July 12th, 1991 to May 19th, 1998", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "12 months",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(0,10) + 
  theme_minimal()

```

```{r}
#| label: fig-linegraphPRKS
#| fig-cap: Changes in United Parks and Recreation Stock Price from One Week Before Blackfish was Released to One Week After Their Pledge to End Orca Breeding
#| echo: false

cleaned_data_PRKS |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "July 12th, 2013 to May 17th, 2016", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "12 months",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(15,40) + 
  theme_minimal()

```

```{r}
#| label: fig-linegraphSBUX
#| fig-cap: Changes in Starbucks Stock Price from One Week Before Legal Action Against Workers United to March 28th
#| echo: false

cleaned_data_SBUX |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "October 11th, 2023 to March 28th, 2024 ", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "1 month",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(80,110) + 
  theme_minimal()
```

# Results {#sec-results}

## Difference in Stock Price Between Commencement and End of Boycott
@tbl-starttoendNKE, @tbl-starttoendPRKS, and @tbl-starttoendSBUX summarizes the change in stock price between a week before the boycott commenced and its commencement date as well as the difference between the start and end dates of the boycott. The first date is the stock price a week before the stock price and the second date is the day the boycott began.

@tbl-starttoendNKE shows Nike had a small decrease in stock price from a week before the beginning of the boycott. There is also an increase in stock price from the day the boycott started to the day it ended. Relative to the stock price, the change is very large as it increased by 335.82%. @tbl-starttoendPRKS shows that there was a significant decrease of 55.3% in United Parks and Recreation's stock price from the beginning to end of the boycott. Starbucks was relatively unchanged from the start to the end of the boycott with a decrease of only 2.5% as shown in @tbl-starttoendSBUX.

```{r}
#| label: tbl-starttoendNKE
#| tbl-cap: Changes in Starbucks Stock Price from One Week Before Legal Action Against Workers United, the the Day it was Announced, to March 28th
#| echo: false

starttoend_NKE <- cleaned_data_NKE |>
  filter(Date == '1991-07-12' | Date == '1991-07-18' | Date == '1998-05-19')

starttoend_NKE <-
  starttoend_NKE |> 
  mutate(Difference = Close - lag(Close))

kable(starttoend_NKE, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference Between Stock Prices"),
      booktabs = TRUE,
      linesep = ""
)
```

```{r}
#| label: tbl-starttoendPRKS
#| tbl-cap: Changes in United Parks & Recreation Stock Price from One Week Before Blackfish was Released, to the Date it was Released, to their Announcement that they will end their Orca Breeding Program
#| echo: false

starttoendPRKS <- cleaned_data_PRKS |>
  filter(Date == '2013-07-12' | Date == '2013-07-19' | Date == '2016-03-16')

starttoendPRKS <-
  starttoendPRKS |> 
  mutate(Difference = Close - lag(Close))

kable(starttoendPRKS, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference Between Stock Prices"),
      booktabs = TRUE,
      linesep = ""
)
```

```{r}
#| label: tbl-starttoendSBUX
#| tbl-cap: Changes in Starbucks Stock Price from One Week Before Legal Action Against Workers United, the the Day it was Announced, to March 28th
#| echo: false

starttoendSBUX <- cleaned_data_SBUX |>
  filter(Date == '2023-10-11' | Date == '2023-10-18' | Date == '2024-03-28')

starttoendSBUX <-
  starttoendSBUX |> 
  mutate(Difference = Close - lag(Close))

kable(starttoendSBUX, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference Between Stock Prices"),
      booktabs = TRUE,
      linesep = ""
)
```

## Detailed Observation of Change in Stock Price
@tbl-twoweeksSBUX shows the daily difference in stock price for two weeks from the announcement of the lawsuit. The stock prices were relatively unaffected immediately following the announcement and November 2nd is responsible for the large spike in price shown in @fig-linegraphSBUX

```{r}
#| label: tbl-twoweeksSBUX
#| tbl-cap: Changes in Starbucks Stock Price Two Weeks From the Lawsuit
#| echo: false

twoweeksSBUX <- 
  cleaned_data_SBUX |>
  filter(between(Date, as.Date('2023-10-18'), as.Date('2023-11-06')))

twoweeksSBUX <-
  twoweeksSBUX |> 
  mutate(Difference = Close - lag(Close))

kable(twoweeksSBUX, 
      col.names = c("Date", 
                    "Closing Stock Price",
                    "Difference"),
      booktabs = TRUE,
      linesep = ""
)
```

## Comparing with the Market
@fig-DJIandNIKE compares the stock price of Nike and the Dow Jones Index (DJI) [@dowjones]. Dow Jones Index includes Nike among around 30 other companies, and it is clear that they have very similar trajectories throughout the boycott. The only apparent difference is the bump in Nike’s price between 1996 and 1998. Otherwise, both stocks have a steady upward growth.

@fig-SPXandPRKS compares the stock price of United Parks & Recreation with the S&P 500 Index (SPX). The comparison cannot indicate whether United Parks was doing better than the market based on price, but by comparing the change in price it is clear that overall the market and United Parks were on a fall entering into 2015. However, the steep price drop at United Parks cannot be explained by the market. The gradual decline at the beginning of the boycott also does not resemble the upward trajectory in the market in 2014.

@fig-ndxSBUX compares Starbucks’ stock price with the Nasdaq 100 Index (NDX). The absence of Starbucks’ price spike in the Nasdaq stock shows that it cannot be attributed to market changes. The following decrease in the Starbucks stock price is also inconsistent with Nasdaq’s upwards trend.

```{r}
#| label: fig-spxPRKS
#| fig-cap: Changes in United Parks and Recreation Stock Price in Comparison with the S&P 500 index
#| echo: false

PRKS_SPX <-
  cleaned_data_SPX |>
  filter(between(date, as.Date('2013-07-19'), as.Date('2016-03-16')))

PRKS_Focus <-
  cleaned_data_PRKS |>
  filter(between(Date, as.Date('2013-07-19'), as.Date('2016-03-16')))

SPX_Stocks <- PRKS_SPX$close
PRKS_Stocks <- PRKS_Focus$Close

PRKS_Date <- PRKS_Focus$Date

# Plot PRKS
plot_prks <- ggplot(PRKS_Focus, aes(x = PRKS_Date, y = PRKS_Stocks)) +
  geom_line(color = "#0000c2") +
  labs(title = "United Parks and Rec Stock Price",
       x = "Date", y = "Stock Price") +
  theme_minimal()

# Plot SPX
plot_spx <- ggplot(PRKS_SPX, aes(x = PRKS_Date, y = SPX_Stocks)) +
  geom_line(color = "#FF5733") +
  labs(title = "S&P 500 Index",
       x = "Date", y = "Index Value") +
  theme_minimal()

# Arrange plots
grid.arrange(plot_prks, plot_spx, ncol = 1)
```

```{r}
#| label: fig-djiNKE
#| fig-cap: Changes in the Nike Stock Price in Comparison with the Dow Jones Index
#| echo: false

NIKE_DJI <-
  cleaned_data_DJI |>
  filter(between(date, as.Date('1991-07-18'), as.Date('1998-05-19')))

NIKE_Focus <-
  cleaned_data_NKE |>
  filter(between(Date, as.Date('1991-07-18'), as.Date('1998-05-19')))

DJI_Stocks <- NIKE_DJI$close
NIKE_Stocks <- NIKE_Focus$Close

NIKE_Date <- NIKE_Focus$Date

# Plot Nike
plot_nike <- ggplot(NIKE_Focus, aes(x = NIKE_Date, y = NIKE_Stocks)) +
  geom_line(color = "#0000c2") +
  labs(title = "Nike Stock Price",
       x = "Date", y = "Stock Price") +
  theme_minimal()

# Plot DJI
plot_dji <- ggplot(NIKE_DJI, aes(x = NIKE_Date, y = DJI_Stocks)) +
  geom_line(color = "#FF5733") +
  labs(title = "Dow Jones Index",
       x = "Date", y = "Index Value") +
  theme_minimal()

# Arrange plots
grid.arrange(plot_nike, plot_dji, ncol = 1)

```

```{r}
#| label: fig-ndxSBUX
#| fig-cap: Changes in the Starbucks Stock Price in Comparison with the Nasdaq 100 Index from October 11th, 2023 to March 28th, 2024
#| echo: false

SBUX_NDX <-
  cleaned_data_NDX |>
  filter(between(date, as.Date('2023-10-11'), as.Date('2024-03-28')))

SBUX_Focus <-
  cleaned_data_SBUX |>
  filter(between(Date, as.Date('2023-10-11'), as.Date('2024-03-18')))

NDX_Stocks <- SBUX_NDX$close
SBUX_Stocks <- SBUX_Focus$Close

SBUX_Date <- SBUX_Focus$Date

# Plot SBUX
plot_SBUX <- ggplot(SBUX_Focus, aes(x = SBUX_Date, y = SBUX_Stocks)) +
  geom_line(color = "#0000c2") +
  labs(title = "SBUX Stock Price",
       x = "Date", y = "Stock Price") +
  theme_minimal()

# Plot NDX
plot_NDX <- ggplot(SBUX_NDX, aes(x = SBUX_Date, y = NDX_Stocks)) +
  geom_line(color = "#FF5733") +
  labs(title = "Nasdaq 100 Index",
       x = "Date", y = "Index Value") +
  theme_minimal()

# Arrange plots
grid.arrange(plot_SBUX, plot_NDX, ncol = 1)
```

# Discussion {#sec-discussion}

##

On November 2nd, Starbucks released their Q4 and Full Year Fiscal Results. Some highlights were record net revenues, store sales up by 8% globally, and record number of stores [@investstarbucks]. In @tbl-twoweeksSBUX, the largest difference in stock price happens on November 2nd at a jump in $8.67. After,

## Out of Sight, Out of Mind: Nike isn't the Only One

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## SeaWorld: The Tainted Consumer Experience

## Starbucks: Impact of Unionization on Stockprice

Investors care about the company's cash flows/earnings margin, when unions are created, the employees get better treatment the earnings margin is affects and is lowered, which investors do not like. Thus, it is likely that stock price went up when Starbucks and Workers United decided to sue each other. They think that Starbucks is gonna win the lawsuit,

On November 2nd, Starbucks released their Q4 and Full Year Fiscal Results. Some highlights were record net revenues, store sales up by 8% globally, and record number of stores [@investstarbucks]

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References